---
title: "Optimal Parameter Search for the Two-Stage Design"
output: html_document
---

# Overview

This folder contains the R functions used to search for optimal design parameters for the proposed two-stage treatment-selection procedure with known common variance and known reference mean.

The main R file is:

`R functions for optimal parameters search.R`

Before running any simulation or optimization, first run this file so that all required R functions are defined in the current R session.

```{r setup, message=FALSE, warning=FALSE}
source("C:/Users/n01654025/Downloads/R functions/R functions for optimal parameters search.R")
```

If the working directory is already the `R functions` folder, use:

```{r eval=FALSE}
source("R functions for optimal parameters search.R")
```

The file defines three helper functions and two main optimization functions.

# Functions

## `stage1_power()`

This helper function computes the Stage-I power under the least favorable configuration.

```r
stage1_power(n1, gamma, delta1, delta2, k, sigma)
```

Arguments:

- `n1`: Stage-I sample size for each experimental treatment.
- `gamma`: Stage-I screening cutoff increment.
- `delta1`: Lower efficacy margin.
- `delta2`: Target efficacy margin.
- `k`: Number of experimental treatments.
- `sigma`: Known common standard deviation.

## `stage2_sample_size()`

This helper function computes the Stage-II sample size required to achieve a specified Stage-II power.

```r
stage2_sample_size(power2, alpha, delta2, sigma)
```

Arguments:

- `power2`: Desired Stage-II power, \(1-\beta_2\).
- `alpha`: One-sided Stage-II significance level.
- `delta2`: Target efficacy margin.
- `sigma`: Known common standard deviation.

## `sampling_quantities()`

This helper function computes the sampling characteristics of a candidate design.

```r
sampling_quantities(n1, n2, gamma, delta1, delta2, k, sigma)
```

Arguments:

- `n1`: Stage-I sample size per experimental treatment.
- `n2`: Stage-II sample size for the selected treatment.
- `gamma`: Stage-I screening cutoff increment.
- `delta1`: Lower efficacy margin.
- `delta2`: Target efficacy margin.
- `k`: Number of experimental treatments.
- `sigma`: Known common standard deviation.

The function returns the probability of terminating after Stage I under the global null, expected sample sizes under the global null and LFC, the sampling-cost criterion, and the maximum possible sample size.

## `optimize_design1()`

Use this function when only the overall power is specified.

```r
optimize_design1(
  n1_range, gamma_range, overall_power, alpha,
  delta1, delta2, k, sigma
)
```

Arguments:

- `n1_range`: Candidate integer values of the Stage-I sample size.
- `gamma_range`: Candidate values of the Stage-I screening cutoff increment.
- `overall_power`: Target overall power, \(1-\beta\).
- `alpha`: One-sided Stage-II significance level.
- `delta1`: Lower efficacy margin.
- `delta2`: Target efficacy margin.
- `k`: Number of experimental treatments.
- `sigma`: Known common standard deviation.

For each candidate `(n1, gamma)` pair, the function computes Stage-I power and retains pairs for which Stage-I power is strictly larger than the target overall power. It then determines the required Stage-II power, computes `n2`, and selects the design minimizing the sampling-cost criterion.

The returned vector contains `n1`, `n2`, `gamma`, `P_terminate_null`, `EN_null`, `EN_LFC`, `cost`, `Nmax`, `power_stage1`, and `power_stage2`.

## `optimize_design2()`

Use this function when both the overall power and Stage-II power are specified.

```r
optimize_design2(
  n1_range, gamma_range, overall_power, stage2_power,
  alpha, delta1, delta2, k, sigma
)
```

Arguments:

- `n1_range`: Candidate integer values of the Stage-I sample size.
- `gamma_range`: Candidate values of the Stage-I screening cutoff increment.
- `overall_power`: Target overall power, \(1-\beta\).
- `stage2_power`: Prespecified Stage-II power, \(1-\beta_2\).
- `alpha`: One-sided Stage-II significance level.
- `delta1`: Lower efficacy margin.
- `delta2`: Target efficacy margin.
- `k`: Number of experimental treatments.
- `sigma`: Known common standard deviation.

The Stage-II power must be strictly larger than the overall power. The function first computes `n2`, then searches over `(n1, gamma)` and retains designs satisfying the Stage-I power requirement. Among feasible designs, it selects the one minimizing the sampling-cost criterion.

The returned vector contains the same quantities as `optimize_design1()`.

# Illustrative Example 1: Table 1

The first example reproduces the simulation used for Table 1.

The simulation settings are:

- \(k=2,3,4,5\);
- overall power \(1-\beta=0.70,0.75,0.80,0.85,0.90\);
- \(\delta_2=0.20,0.30\);
- \(\delta_1=0\);
- \(\alpha=0.05\);
- \(\sigma=1\);
- `n1_range = 10:500`;
- `gamma_range = seq(0, 0.20, by = 0.005)`.

```{r table1-computation}
k_values <- 2:5
overall_power_values <- seq(0.70, 0.90, by = 0.05)
delta2_values <- c(0.20, 0.30)

delta1 <- 0
alpha <- 0.05
sigma <- 1

n1_range <- 10:500
gamma_range <- seq(0, 0.20, by = 0.005)

table_results <- list()
count <- 0

for (k in k_values) {
  for (overall_power in overall_power_values) {
    for (delta2 in delta2_values) {

      fit <- optimize_design1(
        n1_range = n1_range,
        gamma_range = gamma_range,
        overall_power = overall_power,
        alpha = alpha,
        delta1 = delta1,
        delta2 = delta2,
        k = k,
        sigma = sigma
      )

      count <- count + 1

      table_results[[count]] <- data.frame(
        k = k,
        overall_power = overall_power,
        delta2 = delta2,
        n1 = as.integer(fit["n1"]),
        n2 = as.integer(fit["n2"]),
        gamma = as.numeric(fit["gamma"]),
        E0_N = as.numeric(fit["EN_null"]),
        cost = as.numeric(fit["cost"]),
        Nmax = as.integer(fit["Nmax"]),
        one_minus_pi0 = as.numeric(fit["P_terminate_null"]),
        stage1_power = as.numeric(fit["power_stage1"]),
        stage2_power = as.numeric(fit["power_stage2"])
      )
    }
  }
}

design_table1 <- do.call(rbind, table_results)
design_table1 <- unique(design_table1)
rownames(design_table1) <- NULL

if (nrow(design_table1) != 40) {
  warning(paste("Expected 40 rows, but obtained", nrow(design_table1), "rows."))
}
```

## Print Table 1

```{r table1-output}
design_table1
```

```{r table1-kable}
knitr::kable(
  design_table1,
  digits = c(0, 2, 2, 0, 0, 3, 2, 2, 0, 2, 2, 2),
  caption = "Optimal designs when only the overall power is prespecified."
)
```

# Illustrative Example 2: Table 2

The second example uses both prespecified overall power and Stage-II power.

The simulation settings are:

- \(k=2,3,4,5\);
- overall power \(1-\beta=0.70,0.80\);
- Stage-II power \(1-\beta_2=0.80,0.90\), subject to Stage-II power being strictly larger than overall power;
- \(\delta_2=0.20,0.30\);
- \(\delta_1=0\);
- \(\alpha=0.05\);
- \(\sigma=1\);
- `n1_range = 10:500`;
- `gamma_range = seq(0, 0.20, by = 0.005)`.

For overall power 0.70, Stage-II powers 0.80 and 0.90 are used. For overall power 0.80, only Stage-II power 0.90 is used.

```{r table2-computation}
k_values <- 2:5
overall_power_values <- c(0.70, 0.80)
delta2_values <- c(0.20, 0.30)

delta1 <- 0
alpha <- 0.05
sigma <- 1

n1_range <- 10:500
gamma_range <- seq(0, 0.20, by = 0.005)

table2_results <- list()
count <- 0

for (k in k_values) {
  for (overall_power in overall_power_values) {

    if (overall_power == 0.70) {
      stage2_power_values <- c(0.80, 0.90)
    }

    if (overall_power == 0.80) {
      stage2_power_values <- c(0.90)
    }

    for (stage2_power in stage2_power_values) {
      for (delta2 in delta2_values) {

        fit <- optimize_design2(
          n1_range = n1_range,
          gamma_range = gamma_range,
          overall_power = overall_power,
          stage2_power = stage2_power,
          alpha = alpha,
          delta1 = delta1,
          delta2 = delta2,
          k = k,
          sigma = sigma
        )

        count <- count + 1

        table2_results[[count]] <- data.frame(
          k = k,
          overall_power = overall_power,
          stage2_power = stage2_power,
          delta2 = delta2,
          n1 = as.integer(fit["n1"]),
          n2 = as.integer(fit["n2"]),
          gamma = as.numeric(fit["gamma"]),
          E0_N = as.numeric(fit["EN_null"]),
          cost = as.numeric(fit["cost"]),
          Nmax = as.integer(fit["Nmax"]),
          one_minus_pi0 = as.numeric(fit["P_terminate_null"]),
          stage1_power = as.numeric(fit["power_stage1"])
        )
      }
    }
  }
}

design_table2 <- do.call(rbind, table2_results)
design_table2 <- unique(design_table2)
rownames(design_table2) <- NULL

if (nrow(design_table2) != 24) {
  warning(paste("Expected 24 rows, but obtained", nrow(design_table2), "rows."))
}
```

## Print Table 2

```{r table2-output}
design_table2
```

```{r table2-kable}
knitr::kable(
  design_table2,
  digits = c(0, 2, 2, 2, 0, 0, 3, 2, 2, 0, 2, 2),
  caption = "Optimal designs when both overall power and Stage-II power are prespecified."
)
```

# Notes

The optimization uses a grid search over `n1_range` and `gamma_range`. A finer grid can provide a more detailed search but will require more computation time.

If no feasible design is found, enlarge the search ranges.

To generate a GitHub README, knit this file using the `github_document` output format. This creates a `README.md` file that can be uploaded together with the R source file.
